---
title: "Visualizing 3D Genome Organization with gghic"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_document:
    toc: true
    toc_float: true
    toc_depth: 4
    number_sections: true
    theme: cosmo
    highlight: tango
    css: style.css
vignette: >
  %\VignetteIndexEntry{gghic: Visualize Hi-C and Genomic Interactions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
  %\VignetteKeywords{genomics, visualization, Hi-C, 3D genome}
  %\VignetteAbstract{A comprehensive guide to visualizing Hi-C contact maps, TADs, chromatin loops, and other 3D genome features using gghic.}
editor_options:
  markdown:
    wrap: 80
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  tidy = FALSE,
  cache = FALSE,
  dev = "png",
  fig.width = 8,
  fig.height = 6,
  fig.align = "center",
  dpi = 100,
  message = FALSE,
  warning = FALSE,
  fig.alt = "Hi-C visualization"
)
```

## Introduction

**gghic** is an R package for creating publication-ready visualizations of 3D genome organization data. It seamlessly integrates Hi-C contact maps, topologically associating domains (TADs), chromatin loops, gene annotations, and genomic tracks into unified, customizable plots.

### Why gghic?

- **Easy to use**: Simple, intuitive syntax with sensible defaults for quick plotting
- **Flexible**: Choose between high-level `gghic()` wrapper or low-level `geom_*` layers for fine control
- **Memory efficient**: Smart `focus` mechanism loads only the genomic regions you need
- **Publication-ready**: High-quality figures with minimal code
- **Extensible**: Built on `ggplot2` ecosystem—use standard `ggplot2` functions for customization

### Key Features

- **`ChromatinContacts` S4 class**: Robust object-oriented framework for managing Hi-C/-like data
- **Flexible focusing**: Efficiently subset genomic regions before or after loading data
- **Feature integration**: Seamlessly combine TADs, loops, genes, and signal tracks
- **Multiple visualization modes**: Single chromosome, multi-chromosome, inter-chromosomal, and multi-way contacts
- **Hypergraph analysis**: Specialized tools for visualizing Pore-C/-like multi-way chromatin interactions

> **Resources:**
>
> - [arXiv preprint](https://arxiv.org/abs/2412.03005)
> - [GitHub repository](https://github.com/mhjiang97/gghic)
> - [Depth by resolution vignette](depth.html) for Hi-C/-like data quality assessment
> - [Hypergraph vignette](hypergraph.html) for Pore-C/-like multi-way contact analysis

### Typical Workflow

1. **Load data**: Create `ChromatinContacts` object from Cooler files
2. **Focus**: Specify genomic regions of interest (optional but recommended)
3. **Import**: Load interaction data into memory
4. **Add features**: Attach TADs, loops, genes, and tracks (optional)
5. **Visualize**: Use `gghic()` wrapper or custom `ggplot2` layers
6. **Customize**: Refine with standard `ggplot2` functions and themes

---

## Getting Started

### Load Required Packages

```{r setup, message=FALSE, warning=FALSE}
load_pkg <- function(pkgs) {
  for (pkg in pkgs) suppressMessages(require(pkg, character.only = TRUE))
}

load_pkg(
  c("ggplot2", "dplyr", "GenomicRanges", "InteractionSet", "gghic")
)
```

### Download Example data

The example data files are hosted on the [gghic-data](https://github.com/mhjiang97/gghic-data) repository. This function downloads files to a local cache:

```{r message=FALSE, warning=FALSE}
download_example_files <- function(cache_dir, check_exists = TRUE) {
  if (!dir.exists(cache_dir)) dir.create(cache_dir, recursive = TRUE)

  files <- list(
    "chr4_11-100kb.cool" = "cooler/chr4_11-100kb.cool",
    "chr4_11-5kb.cool" = "cooler/chr4_11-5kb.cool",
    "track1.bigWig" = "bigwig/track1.bigWig",
    "track2.bigWig" = "bigwig/track2.bigWig",
    "gencode-chr4_11.gtf.gz" = "gtf/gencode-chr4_11.gtf.gz",
    "TADs_500kb-chr4_11.tsv" = "tad/TADs_500kb-chr4_11.tsv",
    "loops-chr4_11.txt" = "loop/loops-chr4_11.txt",
    "gis_hic.rds" = "multiway/gis_hic.rds",
    "concatemers.rds" = "multiway/concatemers.rds"
  )

  url_base <- "https://raw.githubusercontent.com/mhjiang97/gghic-data/master/"

  for (file_name in names(files)) {
    file_path <- file.path(cache_dir, file_name)
    if (check_exists && file.exists(file_path)) next
    message("Downloading ", file_name, "...")
    download.file(
      paste0(url_base, files[[file_name]]), file_path,
      method = "curl", quiet = TRUE
    )
  }
}
```

```{r message=FALSE, warning=FALSE}
# Define a cache directory and download the files
dir_cache <- file.path("..", "data")
download_example_files(dir_cache)

# Set up paths to the downloaded files
path_cf_100 <- file.path(dir_cache, "chr4_11-100kb.cool")
path_cf_5 <- file.path(dir_cache, "chr4_11-5kb.cool")
path_gtf <- file.path(dir_cache, "gencode-chr4_11.gtf.gz")
paths_track <- file.path(dir_cache, paste0("track", 1:2, ".bigWig"))
path_tad <- file.path(dir_cache, "TADs_500kb-chr4_11.tsv")
path_loop <- file.path(dir_cache, "loops-chr4_11.txt")
path_gis_hic <- file.path(dir_cache, "gis_hic.rds")
path_concatemers <- file.path(dir_cache, "concatemers.rds")
```

## Working with ChromatinContacts Objects

### Understanding ChromatinContacts

The `ChromatinContacts` object is the central data structure in `gghic`. It acts as a lightweight pointer to your Hi-C data (Cooler files), storing resolution and sequence information without loading the massive interaction matrix into memory immediately.

A key feature is the `focus` argument. This allows you to specify chromosomes or regions of interest *before* importing the data.

**Why use focus?**

1. **Speed**: Only reads relevant parts of the Hi-C matrix.
2. **Memory Efficiency**: Prevents loading genome-wide matrices when you only need a specific locus.

### Creating ChromatinContacts Objects

#### No Focus (Genome-Wide)

Without a `focus`, the object references the entire genome:

```{r}
cc_100_all <- ChromatinContacts(cooler_path = path_cf_100)
cc_100_all
```

#### Focusing on Specific Regions

The `focus` argument accepts flexible region specifications. This is the **recommended approach** for working with large Hi-C datasets.

##### Single Region or Chromosome

```{r}
# Focus on entire chromosome (includes all intra-chromosomal interactions)
cc_chr11 <- ChromatinContacts(cooler_path = path_cf_100, focus = "chr11")
cc_chr11
```

```{r}
# Focus on specific 10 Mb region
cc_chr11_region <- ChromatinContacts(
  cooler_path = path_cf_100, focus = "chr11:60000000-70000000"
)
cc_chr11_region
```

##### Multiple Regions with Operators

Use `|` (OR) for multiple regions and `&` (AND) for inter-regional interactions:

```{r}
# Multiple chromosomes
# Returns: chr4-chr4, chr11-chr11, AND chr4-chr11 interactions
cc_multi <- ChromatinContacts(
  cooler_path = path_cf_100, focus = "chr4 | chr11"
)
cc_multi
```

```{r}
# Inter-chromosomal interactions ONLY (no intra-chromosomal)
cc_inter <- ChromatinContacts(
  cooler_path = path_cf_100,
  focus = "chr4:10000000-15000000 & chr11:60000000-65000000"
)
cc_inter
```

##### Complex Focus Patterns

Combine operators for sophisticated queries:

```{r}
# Example 1: Multiple intra-chromosomal + one trans-chromosomal
# Returns: chr1-chr1, chr2-chr2, chr3-chr3, chr1-chr2, chr1-chr3, chr2-chr3, AND chr4-chr5
focus_complex <- c("chr1 | chr2 | chr3", "chr4 & chr5")
```

```{r eval=FALSE}
# Example 2: Specific regions with trans interactions
focus <- c(
  "chr1:1000000-5000000 | chr2:3000000-8000000",
  "chr3:10000000-15000000 & chr4:20000000-25000000"
)

# Example 3: Multiple trans-chromosomal pairs
focus <- c("chr1 & chr2", "chr3 & chr4", "chr5 & chr6")
```

**💡 Practical Tips:**

- Use `focus` when working with genome-wide data files (saves time and memory)
- For exploratory analysis, start with whole chromosome, then zoom to regions
- Use `&` operator when specifically studying trans-chromosomal interactions
- Combine focusing with subsetting (next section) for iterative analysis

##### Using GInteractions Objects

For programmatic region definition (useful in pipelines):

```{r}
# Create a GInteractions object to specify the focus
focus_gi <- suppressWarnings(InteractionSet::GInteractions(
  GenomicRanges::GRanges("chr4:10000000-12000000"),
  GenomicRanges::GRanges("chr11:61000000-63000000")
))

cc_gis_focus <- ChromatinContacts(
  cooler_path = path_cf_100,
  focus = focus_gi
)
cc_gis_focus
```

### Import Interaction Data

After creating the object, use `import()` to load the actual data:

```{r}
cc_100 <- import(cc_100_all)
cc_100
```

```{r}
# Chain operations with pipe operator
cc_5 <- ChromatinContacts(cooler_path = path_cf_5) |>
  import()
cc_5
```

### Adding Genomic Features

Attach genomic features (TADs, loops, tracks) to your `ChromatinContacts` object for integrated visualization.

**Key Benefits:**

- **Automatic synchronization**: Features are automatically subsetted when you zoom or change regions
- **Consistent context**: All features stay aligned with the contact map
- **Simplified plotting**: Pass a single object to `gghic()` instead of managing multiple data sources

```{r}
# Load TADs, loops, and tracks from files
tads <- rtracklayer::import(path_tad, format = "bed")
loops <- path_loop |>
  rtracklayer::import(format = "bedpe") |>
  makeGInteractionsFromGRangesPairs()
tracks <- paths_track |>
  purrr::map(rtracklayer::import) |>
  setNames(paste0("track", seq_along(paths_track))) |>
  GenomicRanges::GRangesList()

# Add features to the ChromatinContacts object
features(cc_100, "TADs") <- tads
features(cc_100, "loops") <- loops
features(cc_100, "tracks") <- tracks

cc_100
```

```{r}
features(cc_5, "TADs") <- tads
features(cc_5, "loops") <- loops
features(cc_5, "tracks") <- tracks
```

### Manipulating ChromatinContacts Objects

`ChromatinContacts` objects support flexible subsetting after data import. This is useful for:

- **Iterative exploration**: Start broad, then zoom to interesting regions
- **Comparative analysis**: Extract same region from multiple samples
- **Quality filtering**: Remove low-quality or extreme-valued interactions

**Automatic updates:** Subsetting filters interactions AND updates:

- Associated features (TADs, loops, tracks)
- The `focus` slot (tracks current genomic context)
- Sequence information (`seqinfo`)

#### Subsetting by Genomic Regions (character)

The most intuitive way to subset is by providing a character string representing a genomic region:

```{r}
# Subset to keep only interactions on chromosome 11
cc_chr11 <- cc_100["chr11"]
cc_chr11
```

```{r}
# Subset to a specific 20 Mb region on chromosome 4
cc_chr4_sub <- cc_100["chr4:20000000-40000000"]
cc_chr4_sub
```

#### Subsetting with GRanges

You can use a `GRanges` object to subset interactions where at least one anchor overlaps the given ranges:

```{r}
# Define a region of interest
roi <- GenomicRanges::GRanges("chr11:60000000-70000000")

# Subset the ChromatinContacts object
cc_gr_sub <- cc_100[roi]
cc_gr_sub
```

#### Subsetting with GInteractions

To select a specific set of interactions, you can subset using a `GInteractions` object. This will keep only the interactions from the `ChromatinContacts` object that overlap with the query `GInteractions`:

```{r}
# Define a specific inter-chromosomal interaction to query
query_gi <- InteractionSet::GInteractions(
  GenomicRanges::GRanges("chr11:60000000-65000000"),
  GenomicRanges::GRanges("chr4:10000000-15000000")
)

# Subset the object
cc_gi_sub <- cc_100[query_gi]

cc_gi_sub
```

#### Subsetting by Index or Logical Vector

You can also use numeric indices or logical vectors for subsetting, similar to how you would subset a `data.frame` or a vector:

```{r}
# Subset by numeric index (e.g., first 1000 interactions)
cc_numeric_sub <- cc_100[1:100]

cc_numeric_sub
```

```{r}
# Subset using a logical vector (e.g., interactions with a score > -0.1)
df <- scaleData(cc_100, "balanced", log10)

# Then subset based on the score
logical_vec <- (df$score > -0.1) & !is.na(df$score)
cc_logical_sub <- cc_100[logical_vec]

cc_logical_sub
```

## Creating Visualizations

`gghic` provides two complementary approaches for plotting:

### Visualization Approaches

#### 1. High-Level Wrapper: `gghic()`

**Best for:** Quick plots, standard layouts, exploratory analysis

- Single function call with sensible defaults
- Automatically handles data scaling and transformation
- Optional arguments for ideograms, genes, TADs, loops, and tracks
- Returns a `ggplot2` object for further customization

#### 2. Low-Level Layers: `geom_*()`

**Best for:** Custom layouts, complex multi-panel figures, precise control

- Build plots layer-by-layer like standard `ggplot2`
- Full control over aesthetics, scales, and themes
- Mix with other `ggplot2` geoms and extensions
- Ideal for publication figures with specific requirements

---

### Basic Hi-C Heatmap

#### Using `geom_hic()` Layer

`geom_hic()` creates the triangle heatmap visualization. It works with standard data frames/tibbles:

**Required aesthetics:**
- `seqnames1`, `start1`, `end1`: First genomic anchor
- `seqnames2`, `start2`, `end2`: Second genomic anchor
- `fill`: Interaction strength (typically log-transformed)

```{r}
df_100 <- cc_100 |>
  scaleData("balanced", log10)

p <- df_100 |>
  dplyr::filter(seqnames1 == "chr11", seqnames2 == "chr11") |>
  ggplot2::ggplot(
    ggplot2::aes(
      seqnames1 = seqnames1, start1 = start1, end1 = end1,
      seqnames2 = seqnames2, start2 = start2, end2 = end2, fill = score
    )
  ) +
  geom_hic()

p
```

`theme_hic()` provides a clean, publication-ready appearance:

- Removes axis labels and tick marks
- Eliminates grid lines
- Focuses attention on the contact map
- Suitable for most Hi-C visualizations

```{r}
p + theme_hic()
```

#### Using `gghic()` Wrapper

The wrapper handles scaling and plotting automatically:

```{r eval=FALSE, echo=TRUE}
cc_100["chr11"] |>
  gghic()
```

### Adding Chromosome Ideograms

Chromosome ideograms provide genomic context by showing:
- The entire chromosome with cytogenetic bands
- Current viewing region highlighted
- Chromosome name and position

#### Using `geom_ideogram()` Layer

**Key parameters:**
- `genome`: Reference genome ("hg19", "hg38", "mm10", "mm39")
- `highlight`: Whether to show the current region as a colored bar
- `length_ratio`: Ideogram height relative to plot (0-1)
- `fontsize`: Text size for chromosome labels

```{r}
cc_100["chr4:10000000-15000000"] |>
  gghic() +
  geom_ideogram(
    genome = "hg19", highlight = TRUE, length_ratio = 0.7, fontsize = 8
  )
```

#### Using the `ideogram` Argument in `gghic()`

```{r eval=FALSE, echo=TRUE}
cc_100["chr4:10000000-15000000"] |>
  gghic(
    ideogram = TRUE, genome = "hg19", highlight = TRUE,
    length_ratio = 0.7, ideogram_fontsize = 8
  )
```

### Adding Gene Annotations

Visualize gene structures to correlate chromatin interactions with genomic features.

**Annotation features:**
- Exon blocks (thick rectangles)
- Introns (connecting lines)
- Gene directionality (arrows for `style = "arrow"`)
- Multiple gene tracks for overlapping genes
- Automatic collision detection and layout

#### Using `geom_annotation()` Layer

```{r}
# Prepare data at 5kb resolution
df_5 <- cc_5 |>
  scaleData("balanced", log10)

# Create a base plot for a small region
p <- df_5 |>
  dplyr::filter(
    seqnames1 == "chr11", seqnames2 == "chr11",
    start1 > 67000000, start1 < 67100000,
    start2 > 67000000, start2 < 67100000
  ) |>
  ggplot(aes(
    seqnames1 = seqnames1, start1 = start1, end1 = end1,
    seqnames2 = seqnames2, start2 = start2, end2 = end2, fill = score
  )) +
  geom_hic() +
  theme_hic()

# Add the annotation layer
p + geom_annotation(gtf_path = path_gtf, style = "basic", maxgap = 100000)
```

**Key parameters:**

- `gtf_path`: Path to GTF/GFF annotation file
- `style`:
  - `"basic"`: Simple exon-intron structure
  - `"arrow"`: Adds directional arrows showing transcription direction
- `maxgap`:
  - Positive integer: Maximum gap between gene tracks (in bp)
  - `-1`: Force all genes on single track (may overlap)
  - `100000`: Good default for most regions
- `gene_symbols`: Character vector of specific genes to display (others hidden)
- `include_ncrna`: Include non-coding RNAs (lncRNA, miRNA, etc.)

```{r}
p + geom_annotation(
  gtf_path = path_gtf, style = "arrow", maxgap = -1,
  gene_symbols = c("GRK2", "SSH3", "KDM2A")
)
```

#### Using the `annotation` Argument in `gghic()`

Again, you can also use the `annotation` argument in `gghic()` to add gene annotations directly.

```{r eval=FALSE, echo=TRUE}
cc_5["chr11:67000000-67100000"] |>
  gghic(
    ideogram = TRUE, genome = "hg19", highlight = FALSE, length_ratio = 0.7, ideogram_fontsize = 8,
    annotation = TRUE, gtf_path = path_gtf, style = "arrow", maxgap = -1, gene_symbols = c("GRK2", "SSH3", "KDM2A")
  )
```

### Adding TADs (Topologically Associating Domains)

TADs are visualized as triangles overlaid on the contact map, representing self-interacting genomic regions.

**Visualization features:**
- Triangle apex points to TAD boundary
- Triangle sides span the TAD region
- Color and line style customizable
- Automatically scaled to match heatmap coordinates

**Two geoms available:**
- `geom_tad()`: Reads TAD file directly (convenient)
- `geom_tad2()`: Uses data frame (more flexible for customization)

#### Using `tad` Argument in `gghic()`

```{r}
cc_100["chr4:50000000-80000000"] |>
  gghic(
    tad = TRUE, tad_is_0based = TRUE, tad_path = path_tad,
    tad_colour = "#00ff83"
  )
```

#### Using `geom_tad()` and `geom_tad2()`

Using `geom_tad2()` makes it easy to customize TAD aesthetics since you can provide a `data.frame` with additional columns.

```{r}
df_tad <- path_tad |>
  read.table(
    sep = "\t", header = FALSE, col.names = c("seqnames", "start", "end")
  ) |>
  dplyr::mutate(start = start + 1) |>
  dplyr::filter(seqnames == "chr4", start > 60000000, end < 70000000) |>
  dplyr::mutate(sample = c("A", "B"))

df_100 |>
  dplyr::filter(
    seqnames1 == "chr4", seqnames2 == "chr4",
    start1 > 60000000, end1 < 70000000,
    start2 > 60000000, end2 < 70000000
  ) |>
  gghic(scale_column = "score", scale_method = function(x) x) +
  geom_tad2(
    data = df_tad, ggplot2::aes(
      seqnames = seqnames, start = start, end = end, colour = sample
    ), stroke = 2
  ) +
  ggplot2::scale_color_grey()
```

### Adding Chromatin Loops

Chromatin loops represent long-range interactions between specific genomic loci (e.g., enhancer-promoter contacts).

**Visualization:**
- Drawn as arcs/semicircles connecting loop anchors
- Arc height proportional to genomic distance
- Customizable colors, line widths, and styles

**Two geoms available:**
- `geom_loop()`: Reads BEDPE file directly
- `geom_loop2()`: Uses data frame with custom aesthetics (color, size by significance)

#### Using `loop` Argument in `gghic()`

```{r}
cc_5["chr11:61915000-65000000"] |>
  gghic(loop = TRUE)
```

#### Using `geom_loop()` and `geom_loop2()`

Using `geom_loop2()` makes it easy to customize loop aesthetics since you can provide a `data.frame` with additional columns.

```{r}
df_loop <- path_loop |>
  read.table(
    sep = "\t", col.names = c(
      "seqnames1", "start1", "end1", "seqnames2", "start2", "end2"
    )
  ) |>
  dplyr::filter(
    seqnames1 == "chr11", seqnames2 == "chr11",
    start1 > 61925000, end1 < 67480000,
    start2 > 61925000, end2 < 67480000
  )

keep <- sample(nrow(df_loop), 6)
df_loop <- df_loop |>
  dplyr::slice(keep) |>
  dplyr::mutate(
    sample = c(rep("A", 3), rep("B", 3))
  )

cc_5["chr11:61915000-65000000"] |>
  gghic() +
  geom_loop2(
    data = df_loop, ggplot2::aes(
      seqnames1 = seqnames1, start1 = start1, end1 = end1,
      seqnames2 = seqnames2, start2 = start2, end2 = end2, colour = sample
    ), stroke = 1, style = "arc"
  )
```

### Multi-Way Contact Visualization

Pore-C and similar technologies capture multi-way contacts where a single DNA molecule contacts 3+ genomic loci simultaneously.

**What are concatemers?**
- Long DNA molecules captured by nanopore sequencing
- Each molecule contains multiple chromatin contact points
- Represented as sets of genomic intervals with a shared read ID

**Visualization approach:**
- Each read shown as a colored horizontal bar
- Bar width indicates genomic span of contacts
- Overlaid on Hi-C heatmap for context

> 💡 **See also:** [Hypergraph vignette](hypergraph.html) for network-based analysis of multi-way contacts

#### Using `geom_concatemer()` and `geom_concatemer2()`

```{r}
# Load the GInteractions object representing multi-way contacts
gis <- path_gis_hic |>
  readRDS()
df <- scaleData(gis, "balanced", log10)

# Select a few reads to visualize
names_read <- c(
  "0c32bbce-24f8-4006-a75c-0e3504648971",
  "10ae8263-cebc-4cbc-94c1-2c1dc5f22101",
  "ced078ca-2d68-5217-a20e-7dc37dd2d786",
  "06063b40-0271-4f4e-993d-2f202028b697",
  "08bb8fe4-c54b-4a11-b9a0-c9eed2d18933"
)
grs_concatemer <- path_concatemers |>
  readRDS() |>
  GenomicRanges::sort()
grs_concatemer <- grs_concatemer[
  GenomicRanges::seqnames(grs_concatemer) %in% c("chr22")
]
df_concatemers <- grs_concatemer |>
  tibble::as_tibble() |>
  dplyr::filter(read_name %in% names_read)

gghic(gis) +
  geom_concatemer2(
    data = df_concatemers, ggplot2::aes(
      seqnames = seqnames, start = start, end = end, read_group = read_name,
      colour = read_name
    ), check_concatemers = TRUE, width_ratio = 0.03
  ) +
  ggplot2::theme(
    legend.key.size = ggplot2::unit(2, "mm"),
    legend.text = ggplot2::element_text(size = 4)
  )
```

#### Converting Concatemers to GInteractions

You can also generate a `GInteractions` object from a set of concatemers using `concatemers2Gis()` to visualize the contact frequency of the concatemers.

```{r}
grs_concatemer_sub <- grs_concatemer[
  grs_concatemer$read_name %in% names_read
]

gis_concatemer <- concatemers2Gis(grs_concatemer_sub, bin_size = 100000)

gis_concatemer |>
  sort() |>
  tibble::as_tibble() |>
  dplyr::mutate(balanced = count) |>
  gghic(scale_method = function(x) x, ideogram = TRUE) +
  geom_concatemer2(
    data = df_concatemers,
    ggplot2::aes(
      seqnames = seqnames, start = start, end = end, read_group = read_name
    )
  ) +
  ggplot2::theme(
    legend.key.size = ggplot2::unit(2, "mm"),
    legend.text = ggplot2::element_text(size = 4)
  )
```

#### Combining Heatmaps

You can combine two heatmaps using `geom_hic_under()` to visualize both pairwise interactions and multi-way contacts.

```{r}
gghic(gis) +
  geom_concatemer2(
    data = df_concatemers,
    ggplot2::aes(
      seqnames = seqnames, start = start, end = end, read_group = read_name
    )
  ) +
  geom_hic_under(
    data = tibble::as_tibble(gis_concatemer) |> dplyr::mutate(score = count),
    ggplot2::aes(
      seqnames1 = seqnames1, start1 = start1, end1 = end1,
      seqnames2 = seqnames2, start2 = start2, end2 = end2, fill2 = score
    ), draw_boundary = FALSE
  ) |>
  renameGeomAes(new_aes = c("fill" = "fill2")) +
  scale_fill_viridis_c(
    aesthetics = "fill2", option = "A", guide = "legend", name = "count"
  )
```

### Multi-Chromosome Visualization

Visualize interactions across multiple chromosomes in a single unified plot.

**Use cases:**
- Genome-wide interaction patterns
- Trans-chromosomal contacts
- Comparative Hi-C across chromosomes
- Chromosome territory analysis

**Layout features:**
- Each chromosome occupies a diagonal block
- Inter-chromosomal interactions shown in off-diagonal blocks
- Chromosome boundaries clearly demarcated

**Key parameters:**
- `draw_boundary = TRUE`: Draws lines separating chromosomes
- `expand_xaxis = TRUE`: Adds spacing between chromosomes for readability
- `expand_left`: Additional left margin (in bp) when using ideograms/annotations

```{r}
cc_100 |>
  gghic(draw_boundary = TRUE, expand_xaxis = TRUE)
```

Chromosome ideograms and gene tracks can also be added to the plot when visualizing multiple chromosomes.

```{r}
p <- df_5 |>
  dplyr::filter(
    start1 > 10000000 & start1 < 11000000 &
      start2 > 10000000 & start2 < 11000000
  ) |>
  gghic(
    scale_column = "score", scale_method = function(x) x,
    ideogram = TRUE, genome = "hg19", highlight = TRUE, ideogram_fontsize = 7, ideogram_width_ratio = 0.08,
    annotation = TRUE, include_ncrna = FALSE, gtf_path = path_gtf, maxgap = 100000, annotation_fontsize = 5, annotation_width_ratio = 0.05,
    draw_boundary = TRUE, expand_xaxis = TRUE, expand_left = 300000
  )

p
```

### Adding Signal Tracks

Integrate 1D genomic signals (ChIP-seq, ATAC-seq, RNA-seq, etc.) alongside Hi-C data.

**Supported formats:**
- BigWig files (`.bw`, `.bigWig`)
- Automatic data range detection or manual specification
- Multiple tracks with independent colors and scales

**Visualization:**
- Tracks displayed as area/mountain plots
- Positioned on left side of heatmap
- Aligned with genomic coordinates
- Rasterization option for large datasets

#### Using `geom_track()` Layer

```{r}
p + geom_track(
  data_paths = paths_track, width_ratio = 0.3, data_range = "auto",
  fill = c("#DC0000B2", "#00A087B2"), rasterize = TRUE
)
```

#### Using `track` Argument in `gghic()`

`gghic()` is handy for adding all these features to the plot.

```{r}
keep <- which(
  df_5$start1 > 10000000 & df_5$start1 < 11000000 &
    df_5$start2 > 10000000 & df_5$start2 < 11000000
)

cc_5[keep] |>
  gghic(
    draw_boundary = TRUE, rasterize = TRUE,
    ideogram = TRUE, genome = "hg19", highlight = TRUE, ideogram_fontsize = 7, ideogram_width_ratio = 0.08,
    annotation = TRUE, include_ncrna = FALSE, gtf_path = path_gtf,
    annotation_style = "arrow", maxgap = 100000, annotation_fontsize = 5, annotation_width_ratio = 0.05,
    track = TRUE, track_width_ratio = 0.5, track_fill = c("#DC0000B2", "#00A087B2"), data_range = "maximum",
    loop = TRUE, loop_style = "arc", stroke = 0.5,
    expand_xaxis = TRUE, expand_left = 300000
  )
```

## Session Information

```{r}
sessionInfo()
```
